---
title: "Mono Lake Levels"
author: "Carly Norris"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
```

```{r}
# read in data
lake_mono <- read_csv(here("data", "mono_lake_annual_levels.csv"), skip = 5) %>% 
  clean_names()
```

```{r}
# plot data for lake levels
  ggplot(data = lake_mono, aes(x=year, y=lake_level_feet_above_sea_level))+ 
  geom_line()+
  scale_x_continuous(limits = c(1850, 2017),
                     breaks = c(1850,1875,1900, 1925, 1950, 1975, 2000, 2017))+
  labs(x= "Year", y= "Water Level (ft. Above Sea Level)",
       title= "Change in Lake Mono Water Levels (Feet Above Sea Level)",
       subtitle = "(1850-2017)")+
  geom_hline(yintercept = 6377,
             color= "cyan4",
             size= 0.6)+
  geom_text(aes(1875, 6377,
                label= "Land Bridge Emerges (6,377 ft.)",
                vjust= -0.5), size= 4)+
   geom_hline(yintercept = 6377,
             color= "cyan4",
             size= 0.6)+
  geom_text(aes(1875, 6377,
                label= "Land Bridge Emerges (6,377 ft.)",
                vjust= -0.5), size= 4)

  

```
**Figure 1:** ADD CAPTION


Lines to add: 3692 stable level salt concentrations, salty shrimp level

1941- LA began diverting water 
1976- first ecological study lake still supports abundant life (brine shrimp, birds)
1983 - supreme court rules to protect Lake Mono as public trust
1984 - streams begin to flow after water diversions ruled against
1994- Decision 1631 set lake levels to 6392 ft
1998- restoration plans for Mono basin streams to get to 6392 ft
2013 - Mono basin stream restoration agreement submitted to state water board and signed
2014- DWP still preventing stream restoration

geom_h