---
title: "Mono Lake Levels (1850-2017)"
author: "Carly Norris"
output: html_document
---

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(dplyr)
```

```{r}
# read in data
lake_mono <- read_csv(here("data", "mono_lake_annual_levels.csv"), skip = 5) %>% 
  clean_names()

# make new events data frame
events_lake <- lake_mono %>% 
  filter(year %in% c("1941", "1994"))
```

```{r}
# plot data for lake levels
  ggplot(data = lake_mono, aes(x=year, y=lake_level_feet_above_sea_level))+ 
  geom_line(size= .8)+
  scale_x_continuous(limits = c(1850, 2017),
                     breaks = c(1850,1875,1900, 1925, 1950, 1975, 2000, 2017))+
  labs(x= "Year", y= "Water Level (ft. Above Sea Level)",
       title= "Change in Lake Mono Water Levels (Feet Above Sea Level)",
       subtitle = "(1850-2017)")+
  geom_hline(yintercept = 6377,
             color= "cyan4",
             size= 0.8)+
  geom_text(aes(1875, 6377,
                label= "Land Bridge Emerges (6,377 ft.)",
                vjust= -0.5))+
   geom_hline(yintercept = 6392,
             color= "darkorchid3",
             size= 0.8)+
  geom_text(aes(1875, 6392,
                label= "Stable Lake Level (6,392 ft.)",
                vjust= -0.5, hjust= 0.6))+
  geom_point(data = events_lake, aes(x= year, y= lake_level_feet_above_sea_level),
             size= 2.5,
             color= "deeppink3") +
  annotate("text", x=1941, y= 6417.0, label= "LA water diversions begin", vjust= -.5, hjust= -.1, size= 3)+
  annotate("text", x= 1994, y= 6374.6, label= "Decision 1961", hjust = -.1, size=3)+
  theme_light()
  

  

```

**Figure 1:** Changes in Lake Mono Water levels (ft. above sea level) from 1850-2017, with significant water levels and dates highlighted. In 1941, LA Department of Water & Power began diverting water to meet growing water demands, resulting in water levels dropping below ecologically stable levels such as land bridge emergence at 6,377 and high lake salinity. In 1994, Decision 1631 set the lake management level to 6,392 ft. to reduce salinity and maintain a balanced ecosystem. (Note: In order to visualize changes in water levels, the y-axis does not start at zero.)  


---
Citations: 

1850-1912 from Stine, Scott (based on occasional observations and San Francisco precipitation). 1912-1979 from LADWP and USGS compilations. 1979-present from Los Angeles Aqueduct Daily Reports, and observations by the Mono Lake Committee. Compiled by the Mono Lake Committee.

Saving Mono Lake. (2020, October 26). Retrieved December 08, 2020, from https://www.monolake.org/learn/aboutmonolake/savingmonolake/
