---
title: "Political Affiliation vs. Environmentalist Self-Identification"
author: "Carly Norris"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
library(broom)
```

## Introduction

This report examines the relationship between individuals self-identification as an environmentalist (yes, no, unsure) and political affiliation using a binary response of liberal vs. conservative. Data was taken from an anonymous survey conducted by Chelsea Batavia and Andrews Forest in 2017 (n= 1,331), exploring the influences on charitable giving for conservation (see citation). A Chi-square test will be used to understand if variables are independent. 


```{r}
# read in data
conservation_survey_simple <- read_csv(here("data", "conservation_survey.csv")) %>% 
  clean_names() %>% 
  select(env_id, pol_or) # only select political affiliation and environmentalist

# organizing by political alignment into only two values representing conservative (pol_or = 1-3) or liberal (pol_or = 5-7)
survey_binary_party <- conservation_survey_simple %>% 
  filter(pol_or == c(1, 2, 3, 5, 6, 7)) %>% 
  mutate(pol_align = case_when(pol_or <= 3 ~ "Conservative",
                               pol_or >= 5 ~ "Liberal"))
```
---

#### Table 1: Political alignment (conservative or liberal) vs. enivronmentalist self-identification (yes, no, unsure). Counts represented in parenthesis.

```{r}
# table political alignment vs. environmental 
q1_counts <- survey_binary_party %>% 
  tabyl(pol_align, env_id) %>% 
  rename("Political Alignment" = "pol_align", "Yes" = "1", "No" = "2", "Unsure" = "3")

# create proportions table
q1_proportions <- q1_counts %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting(digits = 2) %>% 
  adorn_ns()

# display finalized table
kbl(q1_proportions) %>% 
    kable_material(full_width = F)
```

```{r, include= FALSE}
# Chi-squared test for variable independence between self-identification as an environmentalist and political alignment

# convert data to contingency table
q1_ct <- q1_counts %>% 
  column_to_rownames(var = "Political Alignment")

# run chi-square test
q1_x2 <- chisq.test(q1_ct)

# convert to tidy data for in line referencing
q1_tidy <- tidy(q1_x2)
q1_x2
```

There is a significant association between political alignment and environmentalist self-identification ($\chi$^2^(`r q1_tidy$parameter`) = `r  round(q1_tidy$statistic, 2)`, *p* = `r q1_tidy$p.value`), meaning there is a significant association between political alignment (binary affiliation as either conservative or liberal) and survey respondents self-identification as an environmentalist. `r q1_proportions[2,2]` of "liberal" survey respondents considered themselves environmentalists compared to `r q1_proportions[1,2]` of "conservatives." Conversely, `r q1_proportions[1,3]` of "conservatives" did not consider themselves environmentalists vs. `r q1_proportions[2,3]` of "liberals." `r q1_proportions[1,4]` of "conservatives" and `r q1_proportions[2,4]` of "liberals" were unsure if they were environmentalists. 

---
Citation: Andrews Forest LTER Site and C. Batavia. 2019. Influences on charitable giving for conservation: Online survey data of 1,331 respondents across the US, August 2017 ver 3. Environmental Data Initiative. https://doi.org/10.6073/pasta/926e6270e324a1322a900da14d38b96c

